Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2016-11-26 Master Thesis/synopsis.tex
Clarify inequality in 4.2-pre
Clarify isomorphism in 4.2

Now we drop the assumption that $k/\mathbb{Q}$ is Galois. There exists a finite extension $k_1/k$ such that $k_1/\mathbb{Q}$ is Galois and $X(k_1)\neq\varnothing$. Let $K_1 = Kk_1$ be the constant field extension of $K$ with $k_1$ and let $k'_1$ be the constant field of $K'_1$. Applying the above argument to $K_1$ and $K'_1$ yields $[k_1:\mathbb{Q}] = [k'_1:\mathbb{Q}]$. Therefore:
\begin{align*}
[k:\mathbb{Q}] &= [k_1:\mathbb{Q}]\cdot [K_1:K]^{-1} = [k'_1:\mathbb{Q}]\cdot [K'_1:K']^{-1}\\
&\leq [k'_1:\mathbb{Q}]\cdot [k'_1:k']^{-1} = [k':\mathbb{Q}]
\end{align*}
By symmetry, we get $[k:\mathbb{Q}] = [k':\mathbb{Q}]$.
\end{proof}

\begin{lemma}\label{4.2}
We have $(K\overline{k})'=K'\overline{k'}$, i.e. $\sigma(G_{K\overline{k}})=G_{K'\overline{k'}}$. $\sigma$ induces an isomorphism $G_k\to G_{k'}$ such that the following diagram commutes:
\[ \begin{tikzcd}
G_K \ar[r, "\sigma"] \ar[d, "\text{res}"'] & G_{K'}\ar[d, "\text{res}"]\\
G_k \ar[r, "\sigma"'] & G_{k'}
\end{tikzcd} \]
Moreover, $\sigma:G_k\to G_{k'}$ is induced by a unique isomorphism $\varphi:\overline{k'}\to\overline{k}$.
\end{lemma}

\begin{proof}
Let $L/K$ be a finite subextension of $\overline{K}/K$. Let $\ell$ and $\ell'$ be the constant fields of $L$ and $L'$ respectively. Applying Lemma \ref{4.2-pre} to $K$ and $L$ gives:
\[ [\ell:k] = [\ell':k'] \]
Therefore $L/K$ is a constant field extension iff $L'/K'$ is a constant field extension. Since $K\overline{k}/K$ and $K'\overline{k'}/K$ are the union of their finite subextensions, we obtain $(K\overline{k})' = K'\overline{k'}$. This implies that $\sigma$ induces an isomorphism $G_k\to G_{k'}$. Everything else follows from Neukirch-Uchida \ref{thm:neukirch-uchida}.
\end{proof}

>>>>

Now we drop the assumption that $k/\mathbb{Q}$ is Galois. There exists a finite extension $k_1/k$ such that $k_1/\mathbb{Q}$ is Galois and $X(k_1)\neq\varnothing$. Let $K_1 = Kk_1$ be the constant field extension of $K$ with $k_1$ and let $k'_1$ be the constant field of $K'_1$. Applying the above argument to $K_1$ and $K'_1$ yields $[k_1:\mathbb{Q}] = [k'_1:\mathbb{Q}]$. Therefore:
\begin{align*}
[k:\mathbb{Q}] &= [k_1:\mathbb{Q}]\cdot [K_1:K]^{-1} = [k'_1:\mathbb{Q}]\cdot [K'_1:K']^{-1}\\
&= [k'_1:\mathbb{Q}]\cdot ([K'_1: K'k_1'] \cdot [K'k_1':K'])^{-1}\\
&= [k'_1:\mathbb{Q}]\cdot  ([K'_1: K'k_1'] \cdot [k'_1:k'])^{-1} \leq [k'_1:\mathbb{Q}]\cdot [k'_1:k']^{-1} = [k':\mathbb{Q}]
\end{align*}
By symmetry, we get $[k:\mathbb{Q}] = [k':\mathbb{Q}]$.
\end{proof}

\begin{lemma}\label{4.2}
We have $(K\overline{k})'=K'\overline{k'}$, i.e. $\sigma(G_{K\overline{k}})=G_{K'\overline{k'}}$. $\sigma$ induces an isomorphism $G_k\to G_{k'}$ such that the following diagram commutes:
\[ \begin{tikzcd}
G_K \ar[r, "\sigma"] \ar[d] & G_{K'}\ar[d]\\
G_k \ar[r, "\sigma"'] & G_{k'}
\end{tikzcd} \]
Moreover, $\sigma:G_k\to G_{k'}$ is induced by a unique isomorphism $\varphi:\overline{k'}\to\overline{k}$.
\end{lemma}

\begin{proof}
Let $L/K$ be a finite subextension of $\overline{K}/K$. Let $\ell$ and $\ell'$ be the constant fields of $L$ and $L'$ respectively. Applying Lemma \ref{4.2-pre} to $K$ and $L$ gives:
\[ [\ell:k] = [\ell':k'] \]
Therefore $L/K$ is a constant field extension iff $L'/K'$ is a constant field extension. Since $K\overline{k}/K$ and $K'\overline{k'}/K$ are the union of their finite subextensions, we obtain $(K\overline{k})' = K'\overline{k'}$. This implies that $\sigma$ induces an isomorphism:
\[G_k \cong \operatorname{Gal}(K\overline{k}/K) \to \operatorname{Gal}(K'\overline{k'}/K')\cong G_{k'}\] 
Everything else follows from Neukirch-Uchida \ref{thm:neukirch-uchida}.
\end{proof}
